### 2021-01-12 paper/main.tex
soledad final edits

% to-do
% -----
% - To be submitted to EPJ Data Science https://epjdatascience.springeropen.com/about
% - Audit for use of \, as the multiply operator?
% - HOGG: List clearly the assumptions underlying CVMSE and jackknife. They make assumptions about independence of the data points and also something maybe about per-point information content.
% - SOLEDAD: address the point that some constraints may not be satisfiable
% - Soledad: Once appendix is done, point down to it from the text at relevant places.

@@

\section*{\raggedright Fitting very flexible models: Linear regression with large numbers of parameters%
\footnote{%
It is a pleasure to thank
Jed Brown (CU Boulder),
Dan Foreman-Mackey (Flatiron),
Alessandro Gentilini,
Teresa Huang (JHU),
Sam Roweis (deceased),
Adrian Price-Whelan (Flatiron),
Bernhard Sch\"olkopf (MPI-IS),
Kate Storey-Fisher (NYU),
Rachel Ward (UT Austin), and
Lily Zhao (Yale)
for valuable conversations and input.
SV is partially funded by NSF DMS 2044349, EOARD FA9550-18-
1-7007, and the NSF-Simons Research Collaboration on the Mathematical and Scientific
Foundations of Deep Learning (MoDL) (NSF DMS-2031985).}}

@@

The idea of least-squares fitting is that the ``best'' values of the parameters $\beta_j$ are the values that minimize the sum of squares of the differences between the data and the linear combination of features:
\begin{equation}\label{eq:opt1}
    \hat{\beta} = \arg\min_\beta \|Y - X\,\beta\|_2^2
    ~,
\end{equation}
where $\hat{\beta}$ is the $p$-vector (column vector) of the $p$ best-fit values $\hat{\beta}_j$ of the parameters $\beta_j$, and $\|q\|_2^2$ denotes the squared L2-norm, or the \emph{sum of squares} of the components of $q$
\begin{equation}
    \|q\|_2^2 \equiv q^\top q
    ~.
\end{equation}
This optimization objective \eqref{eq:opt1} is convex and the optimization problem has a solution in closed form\footnote{
Note that $\nabla_{\beta} \|Y-X\beta\|^2 = 2X^\top(Y-X\beta)$ so critical points are such that $X^\top Y= X^\top X\beta$. Since the objective is convex all these critical points are global minima. 
} as long as the number of parameters (and features) $p$ is less than the number of training points $n$ (and the matrix $X^\top X$ is invertible):
\begin{equation}
    \hat{\beta} = (X^\top X)^{-1}\,X^\top Y
    ~.
    \label{OLS-under}
\end{equation}
We will treat the case in which the matrix is not invertible below.
When the investigator knows uncertainties on the training data $y_i$, this expression will change a bit; we begin to discuss that in the next \sectionname.

But recall our setting:
We are using the linear fit to interpolate the data, or de-noise the data, or predict new data.
In these contexts, we don't care about the parameter vector $\hat{\beta}$ itself.
We care only about the predictions at a set of new ``test'' locations $t_\ast$, which will usually be different from the training locations $t_j$.
From the new test times $t_\ast$ we create the test feature matrix $X_\ast$ (by the same feature embedding functions $g_j(t)$).
The prediction $\hat{Y}_\ast$ for the $y$ values at the test locations $t_\ast$ becomes
\begin{equation}\label{eq:OLS}
    \hat{Y}_\ast = X_\ast\,(X^\top X)^{-1}\,X^\top Y
    ~.
\end{equation}
Examples of OLS applied to some toy data\footnote{The algorithm by which the toy data were made---and indeed all of the code executed to make the figures for this \documentname---is available online at \url{https://github.com/davidwhogg/FlexibleLinearModels}.} are shown in \figurename~\ref{fig:ols1}.
This form \eqref{eq:OLS} of the prediction of new test data is called ordinary least squares (OLS). It has many good properties, some of which are encoded in the Gauss--Markov theorem.
In particular, if the noise contributions in the model given in equation \eqref{eq.model} are uncorrelated, have zero mean, and equal variances for $i=1,\ldots, n$ then the Gauss--Markov theorem states that the OLS estimator has the lowest variance within the class of unbiased estimators that are linear in $Y$ (see, for example, \citealt{esl}, Ch 3).
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/OLS-under.pdf}
    \caption{Ordinary least-squares (OLS) fits (continuous lines) to a set of example data points (black dots). Fits are shown for different values of the number of basis functions $p$; there are $n=23$ data points. Here (and in all the \figurename s to follow) we are using the Fourier basis functions in \eqref{eq:basis} and shown in \figurename~\ref{fig:basis}. The data $y_i$ were generated using a function that does not reside in the function space spanned by the basis. Here the $X_\ast$ matrices used in the predictions $\hat{Y}_\ast$ were generated from a fine grid of locations in the location coordinate $t$; the plots of the fine grid of predictions are the continuous lines. The fits with larger $p$ have more flexibility to fit the data than the fit with $p=3$, but the fit at the highest $p$ shows evidence of over-fitting.}
    \label{fig:ols1}
    \end{mdframed}
\end{figure}

@@

It is common to include a regularization that discourages the fit from making use of large amplitudes $\beta_j$.
There are many options, but the simplest is ridge regression (or Tikhonov regularization or L2 regularization), which (in the form that doesn't have data weights) looks like
\begin{equation} \label{eq:ridge}
    \hat{\beta} = \arg\min_\beta \|Y - X\,\beta\|_2^2 + \lambda\,\|\beta\|_2^2
    ~,
\end{equation}
where $\lambda>0$ is a regularization parameter that penalizes large values for elements $\beta_j$ of the parameter vector.
This optimization is also convex.
The ridge-regularized prediction for new data looks like
\begin{equation} \label{eq:ridge_sol}
    \hat{Y}_\ast = X_\ast\,(X^\top X + \lambda\,I)^{-1}\,X^\top Y
    ~,
\end{equation}
where $I$ is the $p\times p$ identity.
In the language of Bayesian inference, $\lambda$ can be seen as the inverse of a prior variance for the parameters $\beta_j$.
The salutary effect of the ridge regularization is shown in \figurename~\ref{fig:ridge}.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/ridge.pdf}
    \caption{Comparison of ordinary least-squares (OLS) and ridge-regression fits (continuous lines) to the example data (black dots). In order to illustrate the differences, we chose the $p=21$-parameter fit, which shows evidence of over-fitting at the edges of the fit range. The regularized fit looks more sensible, though it fits the individual data points less precisely. The choice of regularization parameter $\lambda$ matters; here we used $\lambda=0.1$ (chosen arbitrarily).}
    \label{fig:ridge}
    \end{mdframed}
\end{figure}

@@

We are taught folklore, at a young age, that we can never fit for more parameters than we have data. That is, we can never work at $p>n$.
This isn't true!
Not only is it the case that we \emph{can} work at $p>n$, in many cases we \emph{should} work at $p>n$, and many real-world regressions do.
But it is true that the $p>n$ regime is indeed strange!
In the over-parameterized case, there are many settings of the parameters $\beta_j$ that will literally zero out the differences between the data $Y$ and the linear prediction $X\,\beta$.
The OLS solution, in this case, is defined (somewhat arbitrarily) to be the minimum-norm parameter vector $\beta$ that interpolates the data:
\begin{equation}\label{eq:opt2}
    \hat{\beta} = \arg\min_\beta \|\beta\|_2^2 ~~\mbox{subject to}~~ Y = X\,\beta
    ~.
\end{equation}
Technically this formulation depends on an additional assumption that the feature matrix $X$ is full rank or that the data $Y$ lie in the subspace spanned by $X$.
This is true almost always when $p>n$.
This optimization is again convex and has a unique solution, although that solution will depend on feature weights that we discuss in a moment.
When the investigator knows uncertainties on the training data points $y_i$, this expression will change a bit; we return to that in the next \sectionname.
The under-parameterized and over-parameterized optimization statements \eqref{eq:opt1} and \eqref{eq:opt2} can be unified into one form by considering the limit of light L2 regularization:
\begin{equation}\label{eq:opt3}
    \hat{\beta} = \lim_{\lambda\rightarrow 0^+}\left[\arg\min_\beta \|Y - X\,\beta\|_2^2 + \lambda\,\|\beta\|_2^2\right]
    ~;
\end{equation}
in the limit, this delivers the OLS solution in either case.

@@

The OLS prediction for $p>n$ is \emph{not} affine invariant with respect to $p$-dimensional rotations or rescalings.
That is, rotations and scalings in the feature space \emph{will} affect predictions.
It behooves us to re-scale the features (the $p$ $n$-vectors of the feature matrix $X$) in a sensible way, like for instance, to encourage the fit to use low-frequency features more than high-frequency features (\citealt{xie2020weighted, bah2016sample, rauhut2016interpolation}).
We can encode these feature weights in a $p\times p$ diagonal weight matrix $\Lambda^{-1}$ and the prediction becomes
\begin{equation} \label{eq:weighted_sol}
    \hat{Y}_\ast = X_\ast\,\Lambda^{-1}\,X^\top (X\,\Lambda^{-1}\,X^\top)^{-1}\,Y
\end{equation}
This can be seen as the prediction for new data resulting from the following optimization (again, assuming the data can be interpolated):
\begin{equation} \label{eq:weighted}
    \hat{\beta} = \arg\min_\beta \|\Lambda^{1/2}\,\beta\|_2^2 ~~\mbox{subject to}~~ Y = X\,\beta
    ~,
\end{equation}
which---in analogy to the optimizations \eqref{eq:opt2} and \eqref{eq:opt3}---can also be written as
\begin{equation}
    \hat{\beta} = \lim_{\lambda\rightarrow 0^+}\left[\arg\min_\beta \|Y - X\,\beta\|_2^2 + \lambda\,\|\Lambda^{1/2}\,\beta\|_2^2\right]
    ~.
\end{equation}
This optimization\footnote{If you are a physicist and you don't like the mathematical $\|\cdot\|_2^2$ notation---if you prefer to look at quantities that are obvious scalar forms---then recall that $\|\Lambda^{1/2}\,\beta\|_2^2 = \beta^\top\Lambda\,\beta$.
That right-hand-side object is a gauge-invariant object so long as $\beta$ and $\Lambda$ are also gauge-invariant objects themselves.
Yes, data analysis can have this kind of geometric structure!}
penalizes more strongly the parameters $\beta_j$ corresponding to features $g_j(t)$ with larger values of $[\Lambda^{1/2}]_{jj}$.

@@

In the most general case, you have not just a set of feature weights $\Lambda^{-1}$, you also have a set of data-point weights $C^{-1}$ (which, in standard settings, would be the inverses of the variances of the noise affecting the data points $y_i$).
When you put these all together, the feature-weighted, data-weighted least squares predictions are given by either of these two equivalent expressions:
\begin{align}\label{eq:LambdaC1}
    \hat{Y} &= X_\ast\,(X^\top C^{-1}\,X + \Lambda)^{-1}\,X^\top C^{-1}\,Y
    \\ \label{eq:LambdaC2}
    \hat{Y} &= X_\ast\,\Lambda^{-1}\,X^\top (X\,\Lambda^{-1}\,X^\top + C)^{-1}\,Y
    ~.
\end{align}
We discuss the equivalence of these below in \appendixname~\ref{app:math}. These expressions appear in Bayesian-inference contexts (see, for example, \citealt{products}) when $\Lambda^{-1}$ is the variance of the prior on the parameter vector $\beta$ and $C$ is the variance of the noise on the data vector $Y$.
They are solutions to the optimization
\begin{equation}
    \hat{\beta} = \arg\min_\beta \|C^{-1/2}\,(Y - X\,\beta)\|_2^2 + \|\Lambda^{1/2}\,\beta\|_2^2
    ~.
\end{equation}
The units of $C$ must be the units of $Y$ squared, and the units of $\Lambda^{-1}$ must be the square of the ratio of the units of $Y$ to the units of $X$ (the square of the units of $\beta$). These weighted-feature, weighted-data least-square forms \eqref{eq:LambdaC1} and \eqref{eq:LambdaC2} are good because (if the weightings are chosen appropriately) they lead to smooth solutions that are not required to pass precisely through every data point. This is appropriate in common, real situations in which data are noisy and the world is smooth. Which form you choose, between \eqref{eq:LambdaC1} and \eqref{eq:LambdaC2}, depends on a few things, but primarily $p/n$. If $p<n$ it's both faster and more stable to use \eqref{eq:LambdaC1}; if $p>n$ it's faster and more stable to use \eqref{eq:LambdaC2}. We show a toy example of a fit with both feature weights $\Lambda^{-1}$ and data weights $C^{-1}$ in \figurename~\ref{fig:fwwls}.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/weighted-WLS.pdf}
    \caption{Comparison of the feature-weighted OLS shown in \figurename~\ref{fig:fwols} to the same but also including data weights $C^{-1}$, as in equations \eqref{eq:LambdaC1} and \eqref{eq:LambdaC2}.
    Similarly to \figurename~\ref{fig:wls}, the data weights on the diagonal of $C^{-1}$ are the inverses of the squares of the uncertainties shown as vertical error bars. The feature weights encourage the prediction to be smooth; the data weights permit it to be even smoother because they permit the prediction to miss the training data.
    Prediction results now depend on the specific amplitude or prefactor multiplying the feature weights; a particular value of $0.07\,f^2(\omega)$ was chosen for the diagonal of the feature weight matrix $\Lambda^{-1}$ for this demonstration.}
    \label{fig:fwwls}
    \end{mdframed}
\end{figure}

@@

Given a data set (location--data pairs $t_i, y_i$), what is the best way to empirically estimate, from those data, the out-of-sample prediction error---that is, how do you estimate how well you are likely to predict new data?
The only good answer to this is cross-validation:
In cross-validation, we leave out a part of the data (in the most extreme form, leave out one single data point at a time), train the model using all but the left-out part, and predict the left-out part. Then this process is iterated over all choices of what part (or point) to leave out.
This process is illustrated in \figurename~\ref{fig:loo}, where we show $23$ fits, each of which is trained to the 22 points remaining when we leave one of the $n=23$ points out.
Also shown is the prediction, in each leave-one-out fit, for the left-out point.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/LOO.pdf}
    \caption{A demonstration of leave-one-out cross-validation for one particular model. The 23 lines are the 23 fits, for each of which one data point was held out. Also plotted on each of the 23 lines is the prediction made for that fit's held-out data point. Some of the predictions don't appear within the plot window, because the predictions obtain large amplitudes at the edges.}
    \label{fig:loo}
    \end{mdframed}
\end{figure}

The prediction for the mean-squared error (MSE) for new data is the mean of the square of the differences between the leave-one-out predictions and the left-out data.
This leave-one-out cross-validation MSE (CVMSE) will be different for different choices for the basis, size ($p$), and regularization of the fits you do.
It is a fairly reliable and well-studied method for assessing predictive accuracy (see, for example, \citealt{cv}), although it does depend on some assumptions (for example, about the distribution of predictive information across individual data points, maybe).
An example of CVMSE for two models (the feature-weighted OLS and the ridge regression) are shown for our toy data, as a function of the number of features $p$, in \figurename~\ref{fig:cv}.
The predictive accuracy is indeed a function of $p$.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/cross-validation.pdf}
    \caption{Leave-one-out cross-validation estimates of mean squared prediction error for the feature-weighted OLS fits, as a function of the number of features $p$. Also shown are the same for ridge regression with $\lambda=0.1$ and also the Gaussian process that we will introduce in \sectionname~\ref{sec:gp} and show in \figurename~\ref{fig:gp}; the GP result is shown as a flat line because it doesn't have an associated number $p$. In detail the first (lowest-$t$) and last (highest-$t$) data points were not used in computing the mean squared error; that choice is debatable, but we are imagining an assessment of the quality of the \emph{interpolations}, not extrapolations, of these models. Note that the OLS predictions are much, much worse at $p\approx n$ (check out all the orders of magnitude on the vertical axis) than they are at very low or very high $p$, but that ridge regression doesn't show this behavior.}
    \label{fig:cv}
    \end{mdframed}
\end{figure}

@@

There is one more point to make about over-fitting which is that, when your data set is small, it is possible to over-fit even your cross-validation.
This problem is beyond the scope of this \documentname; all we will say here is that it is not possible to make precise settings of many hyper-parameters through cross validation.
Because in cross validation you are making use of your data to set the properties of your regression, there are dangers of over-adapting your method to your data.
It is safer to have a fully independent validation data set, but this is rarely practical (and it doesn't completely protect you either \citealt{SOLEDAD}).

@@

In the particular case of \eqref{kkt2} it suffices to compute the top right block the inverse. Obtaining that if both $\Lambda$ and $X\Lambda^{-1}X^\top$ are invertible we have 
\begin{equation}
    \hat \beta = \Lambda^{-1}X^\top(X\Lambda^{-1}X^\top)^{-1}Y. 
\end{equation}

>>>>

% This document is part of the FlexibleLinearModels project. Copyright 2021 the authors.

% to-do
% -----
% - To be submitted to EPJ Data Science https://epjdatascience.springeropen.com/about

@@

\section*{\raggedright Fitting very flexible models: Linear regression with large numbers of parameters%
\footnote{%
It is a pleasure to thank
Jed Brown (CU Boulder),
Dan Foreman-Mackey (Flatiron),
Alessandro Gentilini,
Teresa Huang (JHU),
Sam Roweis (deceased),
Adrian Price-Whelan (Flatiron),
Bernhard Sch\"olkopf (MPI-IS),
Kate Storey-Fisher (NYU),
Rachel Ward (UT Austin), and
Lily Zhao (Yale)
for valuable conversations and input.
SV is partially funded by NSF DMS 2044349, EOARD FA9550-18-1-7007, and the NSF--Simons Research Collaboration on the Mathematical and Scientific Foundations of Deep Learning (MoDL) (NSF DMS-2031985).}}

@@

The idea of least-squares fitting is that the ``best'' values of the parameters $\beta_j$ are the values that minimize the sum of squares of the differences between the data and the linear combination of features:
\begin{equation}\label{eq:opt1}
    \hat{\beta} = \arg\min_\beta \|Y - X\,\beta\|_2^2
    ~,
\end{equation}
where $\hat{\beta}$ is the $p$-vector (column vector) of the $p$ best-fit values $\hat{\beta}_j$ of the parameters $\beta_j$, and $\|q\|_2^2$ denotes the squared L2-norm, or the \emph{sum of squares} of the components of $q$
\begin{equation}
    \|q\|_2^2 \equiv q^\top q
    ~.
\end{equation}
This optimization objective \eqref{eq:opt1} is convex and the optimization problem has a solution in closed form\footnote{
Note that $\nabla_{\beta} \|Y-X\beta\|^2 = 2X^\top(Y-X\beta)$ so critical points are such that $X^\top Y= X^\top X\beta$. Since the objective is convex all these critical points are global minima. Further analysis is given in Appendix \ref{app:math}.
} as long as the number of parameters (and features) $p$ is less than the number of training points $n$ (and the matrix $X^\top X$ is invertible):
\begin{equation}
    \hat{\beta} = (X^\top X)^{-1}\,X^\top Y
    ~.
    \label{OLS-under}
\end{equation}
We will treat the case in which the matrix is not invertible below.
When the investigator knows uncertainties on the training data $y_i$, this expression will change a bit; we begin to discuss that in the next \sectionname.

But recall our setting:
We are using the linear fit to interpolate the data, or de-noise the data, or predict new data.
In these contexts, we don't care about the parameter vector $\hat{\beta}$ itself.
We care only about the predictions at a set of new ``test'' locations $t_\ast$, which will usually be different from the training locations $t_j$.
From the new test times $t_\ast$ we create the test feature matrix $X_\ast$ (by the same feature embedding functions $g_j(t)$).
The prediction $\hat{Y}_\ast$ for the $y$ values at the test locations $t_\ast$ becomes
\begin{equation}\label{eq:OLS}
    \hat{Y}_\ast = X_\ast\,(X^\top X)^{-1}\,X^\top Y
    ~.
\end{equation}
Examples of OLS applied to some toy data\footnote{The algorithm by which the toy data were made---and indeed all of the code used to make the figures for this \documentname---is available online at \url{https://github.com/davidwhogg/FlexibleLinearModels}.} are shown in \figurename~\ref{fig:ols1}.
This form \eqref{eq:OLS} of the prediction of new test data is called ordinary least squares (OLS). It has many good properties, some of which are encoded in the Gauss--Markov theorem.
In particular, if the noise contributions in the model given in equation \eqref{eq.model} are uncorrelated, have zero mean, and equal variances for $i=1,\ldots, n$ then the Gauss--Markov theorem states that the OLS estimator has the lowest variance within the class of unbiased estimators that are linear in $Y$ (see, for example, \citealt{esl}, Ch 3).
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/OLS-under.pdf}
    \caption{Ordinary least-squares (OLS) fits (continuous lines) to a set of example data points (black dots). Fits are shown for different values of the number of basis functions $p$; there are $n=23$ data points. Here (and in all the \figurename s to follow) we are using the Fourier basis functions in \eqref{eq:basis} and shown in \figurename~\ref{fig:basis}. The data $y_i$ were generated using a function that does not reside in the function space spanned by the basis. Here the $X_\ast$ matrices used in the predictions $\hat{Y}_\ast$ were generated from a fine grid of locations in the location coordinate $t$; the plots of the fine grid of predictions are the continuous lines. The fits with larger $p$ have more flexibility to fit the data than the fit with $p=3$, but the fit at the highest $p$ shows evidence of over-fitting.}
    \label{fig:ols1}
    \end{mdframed}
\end{figure}

@@

It is common to include a regularization that discourages the fit from making use of large amplitudes $\beta_j$.
There are many options, but the simplest is ridge regression (or Tikhonov regularization or L2 regularization), which (in the form that doesn't have data weights) looks like
\begin{equation} \label{eq:ridge}
    \hat{\beta} = \arg\min_\beta \|Y - X\,\beta\|_2^2 + \lambda\,\|\beta\|_2^2
    ~,
\end{equation}
where $\lambda>0$ is a regularization parameter that penalizes large values for elements $\beta_j$ of the parameter vector.
This optimization is also convex.
The ridge-regularized prediction for new data looks like
\begin{equation} \label{eq:ridge_sol}
    \hat{Y}_\ast = X_\ast\,(X^\top X + \lambda\,I)^{-1}\,X^\top Y
    ~,
\end{equation}
where $I$ is the $p\times p$ identity (see Appendix \ref{app:math}).
In the language of Bayesian inference, $\lambda$ can be seen as the inverse of a prior variance for the parameters $\beta_j$.
The salutary effect of the ridge regularization is shown in \figurename~\ref{fig:ridge}.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/ridge.pdf}
    \caption{Comparison of ordinary least-squares (OLS) and ridge-regression fits (continuous lines) to the example data (black dots). In order to illustrate the differences, we chose the $p=21$-parameter fit, which shows evidence of over-fitting at the edges of the fit range. The regularized fit looks more sensible, though it fits the individual data points less precisely. The choice of regularization parameter $\lambda$ matters; here we used $\lambda=0.1$ (chosen arbitrarily).}
    \label{fig:ridge}
    \end{mdframed}
\end{figure}

@@

We are taught folklore, at a young age, that we can never fit for more parameters than we have data. That is, we can never work at $p>n$.
This isn't true!
Not only is it the case that we \emph{can} work at $p>n$, in many cases we \emph{should} work at $p>n$, and many real-world regressions do.
But it is true that the $p>n$ regime is indeed strange!
In the over-parameterized case, there are many settings of the parameters $\beta_j$ that will literally zero out the differences between the data $Y$ and the linear prediction $X\,\beta$.
The OLS solution, in this case, is defined (somewhat arbitrarily) to be the minimum-norm parameter vector $\beta$ that interpolates the data:
\begin{equation}\label{eq:opt2}
    \hat{\beta} = \arg\min_\beta \|\beta\|_2^2 ~~\mbox{subject to}~~ Y = X\,\beta
    ~.
\end{equation}
Technically this formulation depends on an additional assumption that the feature matrix $X$ is full rank or that the data $Y$ lie in the subspace spanned by $X$.
This is true almost always when $p>n$.
This optimization is again convex and has a unique solution, although that solution will depend on feature weights that we discuss in a moment.
When the investigator knows uncertainties on the training data points $y_i$, this expression will change a bit; we return to that in the next \sectionname.
The under-parameterized and over-parameterized optimization statements \eqref{eq:opt1} and \eqref{eq:opt2} can be unified into one form by considering the limit of light L2 regularization:
\begin{equation}\label{eq:opt3}
    \hat{\beta} = \lim_{\lambda\rightarrow 0^+}\left[\arg\min_\beta \|Y - X\,\beta\|_2^2 + \lambda\,\|\beta\|_2^2\right]
    ~;
\end{equation}
in the limit, this delivers the OLS solution in either case and doesn't require the constraint in \eqref{eq:opt2} to be satisfied.

@@

The OLS prediction for $p>n$ is \emph{not} affine invariant with respect to $p$-dimensional rotations or rescalings.
That is, rotations and scalings in the feature space \emph{will} affect predictions.
It behooves us to re-scale the features (the $p$ $n$-vectors of the feature matrix $X$) in a sensible way, like for instance, to encourage the fit to use low-frequency features more than high-frequency features (\citealt{xie2020weighted, bah2016sample, rauhut2016interpolation}).
We can encode these feature weights in a $p\times p$ diagonal weight matrix $\Lambda^{-1}$ and the prediction becomes
\begin{equation} \label{eq:weighted_sol}
    \hat{Y}_\ast = X_\ast\,\Lambda^{-1}\,X^\top (X\,\Lambda^{-1}\,X^\top)^{-1}\,Y
\end{equation}
This can be seen as the prediction for new data resulting from the following optimization (again, assuming the data can be interpolated):
\begin{equation} \label{eq:weighted}
    \hat{\beta} = \arg\min_\beta \|\Lambda^{1/2}\,\beta\|_2^2 ~~\mbox{subject to}~~ Y = X\,\beta
    ~,
\end{equation}
which---in analogy to the optimizations \eqref{eq:opt2} and \eqref{eq:opt3}---can also be written as
\begin{equation}
    \hat{\beta} = \lim_{\lambda\rightarrow 0^+}\left[\arg\min_\beta \|Y - X\,\beta\|_2^2 + \lambda\,\|\Lambda^{1/2}\,\beta\|_2^2\right]
    ~,
\end{equation}
which doesn't require the constraints in \eqref{eq:weighted} to be satisfied. 
This optimization\footnote{If you are a physicist and you don't like the mathematical $\|\cdot\|_2^2$ notation---if you prefer to look at quantities that are obvious scalar forms---then recall that $\|\Lambda^{1/2}\,\beta\|_2^2 = \beta^\top\Lambda\,\beta$.
That right-hand-side object is a gauge-invariant object so long as $\beta$ and $\Lambda$ are also gauge-invariant objects themselves.
Yes, data analysis can have this kind of geometric structure!}
penalizes more strongly the parameters $\beta_j$ corresponding to features $g_j(t)$ with larger values of $[\Lambda^{1/2}]_{jj}$.

@@

In the most general case, you have not just a set of feature weights $\Lambda^{-1}$, you also have a set of data-point weights $C^{-1}$ (which, in standard settings, would be the inverses of the variances of the noise affecting the data points $y_i$).
When you put these all together, the feature-weighted, data-weighted least squares predictions are given by either of these two equivalent expressions:
\begin{align}\label{eq:LambdaC1}
    \hat{Y} &= X_\ast\,(X^\top C^{-1}\,X + \Lambda)^{-1}\,X^\top C^{-1}\,Y
    \\ \label{eq:LambdaC2}
    \hat{Y} &= X_\ast\,\Lambda^{-1}\,X^\top (X\,\Lambda^{-1}\,X^\top + C)^{-1}\,Y
    ~.
\end{align}
The equivalence of these is due to the Woodbury matrix identity \citep{henderson1981deriving}. These expressions appear in Bayesian-inference contexts (see, for example, \citealt{products}) when $\Lambda^{-1}$ is the variance of the prior on the parameter vector $\beta$ and $C$ is the variance of the noise on the data vector $Y$.
They are solutions to the optimization
\begin{equation}
    \hat{\beta} = \arg\min_\beta \|C^{-1/2}\,(Y - X\,\beta)\|_2^2 + \|\Lambda^{1/2}\,\beta\|_2^2
    ~.
\end{equation}
The units of $C$ must be the units of $Y$ squared, and the units of $\Lambda^{-1}$ must be the square of the ratio of the units of $Y$ to the units of $X$ (the square of the units of $\beta$). These weighted-feature, weighted-data least-square forms \eqref{eq:LambdaC1} and \eqref{eq:LambdaC2} are good because (if the weightings are chosen appropriately) they lead to smooth solutions that are not required to pass precisely through every data point. This is appropriate in common, real situations in which data are noisy and the world is smooth. Which form you choose, between \eqref{eq:LambdaC1} and \eqref{eq:LambdaC2}, depends on a few things, but primarily $p/n$. If $p<n$ it's both faster and more stable to use \eqref{eq:LambdaC1}; if $p>n$ it's faster and more stable to use \eqref{eq:LambdaC2}. We show a toy example of a fit with both feature weights $\Lambda^{-1}$ and data weights $C^{-1}$ in \figurename~\ref{fig:fwwls}.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/weighted-WLS.pdf}
    \caption{Comparison of the feature-weighted OLS shown in \figurename~\ref{fig:fwols} to the same but also including data weights $C^{-1}$, as in equations \eqref{eq:LambdaC1} and \eqref{eq:LambdaC2}.
    Similarly to \figurename~\ref{fig:wls}, the data weights on the diagonal of $C^{-1}$ are the inverses of the squares of the uncertainties shown as vertical error bars. The feature weights encourage the prediction to be smooth; the data weights permit it to be even smoother because they permit the prediction to miss the training data.
    Prediction results now depend on the specific amplitude or prefactor multiplying the feature weights; a particular value of $0.07\,f^2(\omega)$ was chosen for the diagonal of the feature weight matrix $\Lambda^{-1}$ for this demonstration.}
    \label{fig:fwwls}
    \end{mdframed}
\end{figure}

@@

Given a data set (location--data pairs $t_i, y_i$), what is the best way to empirically estimate, from those data, the out-of-sample prediction error?
That is, how do you estimate how well you are likely to predict new data?
A reasonable answer to this is cross-validation:
In cross-validation, we leave out a part of the data (in the most extreme form, leave out one single data point at a time), train the model using all but the left-out part, and predict the left-out part. Then this process is iterated over all choices of what part (or point) to leave out.
This process is illustrated in \figurename~\ref{fig:loo}, where we show $23$ fits, each of which is trained to the 22 points remaining when we leave one of the $n=23$ points out.
Also shown is the prediction, in each leave-one-out fit, for the left-out point.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/LOO.pdf}
    \caption{A demonstration of leave-one-out cross-validation for one particular model. The 23 lines are the 23 fits, for each of which one data point was held out. Also plotted on each of the 23 lines is the prediction made for that fit's held-out data point. Some of the predictions don't appear within the plot window, because the predictions obtain large amplitudes at the edges.}
    \label{fig:loo}
    \end{mdframed}
\end{figure}

The prediction for the mean-squared error (MSE) for new data is the mean of the square of the differences between the leave-one-out predictions and the left-out data.
This leave-one-out cross-validation MSE (CVMSE) will be different for different choices for the basis, size ($p$), and regularization of the fits you do.
It is a fairly reliable and well-studied method for assessing predictive accuracy (see, for example, \citealt{cv}), although it does depend on some assumptions (for example, that data points are not duplicated or strongly corrrelated, and that the predictive information in the data is distributed among multiple individual data points).
An example of CVMSE for two models (the feature-weighted OLS and the ridge regression) are shown for our toy data, as a function of the number of features $p$, in \figurename~\ref{fig:cv}.
The predictive accuracy is indeed a function of $p$.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/cross-validation.pdf}
    \caption{Leave-one-out cross-validation estimates of mean squared prediction error for the feature-weighted OLS fits, as a function of the number of features $p$. Also shown are the same for ridge regression with $\lambda=0.1$ and also the Gaussian process that we will introduce in \sectionname~\ref{sec:gp} and show in \figurename~\ref{fig:gp}; the GP result is shown as a flat line because it doesn't have an associated number $p$. In detail the first (lowest-$t$) and last (highest-$t$) data points were not used in computing the mean squared error; that choice is debatable, but we are imagining an assessment of the quality of the \emph{interpolations}, not extrapolations, of these models. Note that the OLS predictions are much, much worse at $p\approx n$ (check out all the orders of magnitude on the vertical axis) than they are at very low or very high $p$, but that ridge regression doesn't show this behavior.}
    \label{fig:cv}
    \end{mdframed}
\end{figure}

@@

There is one more point to make about over-fitting which is that, when your data set is small, it is possible to over-fit even your cross-validation.
This problem is beyond the scope of this \documentname; all we will say here is that it is not possible to make precise settings of many hyper-parameters through cross validation.
Because in cross validation you are making use of your data to set the properties of your regression, there are dangers of over-adapting your method to your data.
It is safer to have a fully independent validation data set, but this is rarely practical (and it doesn't completely protect you either \citealt{cifar10}).

@@

In the particular case of \eqref{kkt2} it suffices to compute the top right block the inverse. Obtaining that if both $\Lambda$ and $X\,\Lambda^{-1}X^\top$ are invertible we have 
\begin{equation}
    \hat \beta = \Lambda^{-1}\,X^\top(X\,\Lambda^{-1}X^\top)^{-1}\,Y. 
\end{equation}
